---
title: "Simulating Tissues with Oneiric"
author: "Patrick Martin"
output:
    pdf_document: null
    keep_tex: yes
    number_sections: no
    html_document:
        toc: true
package: '`oneiric'''
vignette: |
    %\VignetteIndexEntry{Oneiric User's Guide}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
---

#  Introduction
Benchmarking spatial analysis tools is tricky especially when there exists few ground truth data sets.
Instead, we decided that we would use simulated data since this allows us to have a strong ground truth.
We can also control the parameters used to create the data sets in order to simulate a variety of potential
biological situations. 

We wrapped our simulation in an R package - Oneiric - to faciliate to reproducibility and replicability of our work.


```{r setup, eval = TRUE, echo = TRUE}
library(oneiric)
library(scater)
library(ggplot2)
library(RColorBrewer)
set.seed(288)
```

# Data
Oneiric provides multiples territory structures - one of them being based on a tinkerbell progression 
chaos map. We provide a set of parameters that will produce useable chaos maps. 

It is also possible to search for other parameters using the `find_tinkerbell` function. 

## Chaos Parameters
```{r chaos_param, eval = TRUE, echo  = TRUE}

data(oneiric)
map_params
```

This function will attempt to find parameters for the tinkerbell progression that can be converted into territories. 
Since this is can be a time consuming process, we set a timer for parameter search (in seconds). 
```{r find_tinker, eval = FALSE, echo = TRUE}
map_params <- find_tinkerbell(time = 120,
    n_maps = 96,
    plot = FALSE,
    export = FALSE,
    file_name = "maps")
```

## Preparing output directory

```{r out_dir, eval = TRUE, echo = TRUE}

output <- "/Users/martinp4/Documents/Cedars/Oneiric/simulations/"
```


## Territory Types
Here, we demonstrate the possible territory types that can be created using oneiric. We have four types:

* Circle: randomly size circles placed in a background territory.
* Rod: randomly size rod-like structures placed in a background territory.
* Chaos: tinkerbell progression chaos map placed in a background.
* Layered: any of the above but each territory will divided into layers. 


```{r spatial_sim, eval = TRUE, echo = TRUE}
circle <- simulate_spatial(n_cells = 5000,
    n_territories = 5,
    n_samples = 12,
    pattern = "circle",
    expanse = c(0.1, 0.3))

rod <- simulate_spatial(n_cells = 5000,
    n_territories = 5,
    n_samples = 12,
    pattern = "rod",
    width_range = c(0.0, 0.1),
    length_range = c(0.2, 0.5))

chaos_map <- simulate_spatial(n_cells = 5000,
    n_samples = 12,
    pattern = "chaos",
    expanse = 0.02) # expanse will dilate the territory by a factor of 0.02 (max of 1)

layered_map <- simulate_spatial(n_cells = 5000,
    n_samples = 12,
    n_territories = 1,
    pattern = "circle",
    layers = 5,
    expanse =  c(0.4, 0.5))

```



```{r view_territory_types, eval = TRUE, echo = TRUE}
circles <- do.call("rbind", circle)
circles$Territory <- as.factor(circles$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$Territory)))
g <- ggplot(circles, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


rods <- do.call("rbind", rod)
rods$Territory <- as.factor(rods$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(rods$Territory)))
g1 <- ggplot(rods, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


tinker <- do.call("rbind", chaos_map)
tinker$Territory <- as.factor(tinker$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(tinker$Territory)))
g2 <- ggplot(tinker, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


layers <- do.call("rbind", layered_map)
layers$Territory <- as.factor(layers$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(layers$Territory)))
g3 <- ggplot(layers, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))

```

```{r g_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Circle Territories"}
print(g)
```

```{r g1_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Rod Territories"}
print(g1)
```

```{r g2_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Chaos Map Territories"}
print(g2)
```

```{r g3_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Layered Territories"}
print(g3)
```

# Creating Simulated Data Sets

This section covers how the data sets used for benchmarking Vesalius were produced. 
We produce data sets with no cell labels.

## Simulating Circle regime

```{r circle_solo, eval = TRUE, echo = TRUE}

circular <- simulate_spatial(n_cells = 5000, 
    n_territories = 5, 
    n_samples = 12, 
    pattern = "circle", 
    expanse = c(0.1, 0.25), 
    force_cells = 10)

circular_counts <- simulate_cells(circular,
    cell_composition = 2,
    no_label = TRUE)

circles <- do.call("rbind", circular_counts$spatial)
circles$Territory <- as.factor(circles$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$Territory)))
g <- ggplot(circles, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))



circles$cell_labels <- as.factor(circles$cell_labels)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$cell_labels)))
g1 <- ggplot(circles, aes(x = x, y = y, col = cell_labels)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))

```

```{r no_lab_g_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="No label territories assumes that no cell type information is provided. A single label is given to all cells in the Territory column"}
print(g)
```

```{r no_lab_g1_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="No label territories assumes that no cell type information is provided. Ground truth cell type labels are still provided as the cell_lables column"}
print(g1)
```

```{r export_circle, eval = FALSE, echo =TRUE}
export_simulation(spatial = circular_counts$spatial,
    cells = circular_counts$counts,
    out_dir = output,
    file_tag = "circle_spatial_territories")

```


## Simuating Layer regime

```{r circle_layer_nl, eval = TRUE, echo = TRUE}
circular_layer_nl <- simulate_spatial(n_cells = 5000,
    n_territories = 1,
    n_samples = 12,
    pattern = "circle",
    expanse = c(0.4, 0.5),
    layer = 5,
    force_cells = 12)

circular_layer_counts_nl <- simulate_cells(circular_layer_nl,
    as_layer = TRUE,
    de_prob = 0.5,
    de_layer = 0.05,
    cell_composition = 2,
    no_label = TRUE)

# to mimic no cell type labels available

circles <- do.call("rbind", circular_layer_counts_nl$spatial)
circles$Territory <- as.factor(circles$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$Territory)))
g <- ggplot(circles, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


circles$cell_labels <- as.factor(circles$cell_labels)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$cell_labels)))
g1 <- ggplot(circles, aes(x = x, y = y, col = cell_labels)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))




```
```{r no_lab_lay_g_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="No label territories assumes that no cell type information is provided. A single label is given to all cells in the Territory column"}
print(g)
```

```{r no_lab_lay_g1_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="No label territories assumes that no cell type information is provided. Ground truth cell type labels are still provided as the cell_lables column"}
print(g1)
```

```{r export_layer, eval = FALSE, echo = TRUE}
export_simulation(spatial = circular_layer_counts_nl$spatial,
    cells = circular_layer_counts_nl$counts,
    out_dir = output,
    file_tag = "layered_spatial_territories")
```