---
title: "Simulating Tissues with Oneiric"
author: "Patrick Martin"
output:
    pdf_document: null
    keep_tex: yes
    number_sections: no
    html_document:
        toc: true
package: '`oneiric'''
vignette: |
    %\VignetteIndexEntry{Oneiric User's Guide}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
---

#  Introduction
Benchmarking spatial analysis tools is tricky especially when there exists few ground truth data sets.
Instead, we decided that we would use simulated data since this allows us to have a strong ground truth.
We can also control the parameters used to create the data sets in order to simulate a variety of potential
biological situations. 

We wrapped our simulation in an R package - Oneiric - to faciliate to reproducibility and replicability of our work.


```{r setup, eval = TRUE, echo = TRUE}
library(oneiric)
library(scater)
library(ggplot2)
library(RColorBrewer)
set.seed(288)
```

# Data
Oneiric provides multiples territory structures - one of them being based on a tinkerbell progression 
chaos map. We provide a set of parameters that will produce useable chaos maps. 

It is also possible to search for other parameters using the `find_tinkerbell` function. 

## Chaos Parameters
```{r chaos_param, eval = TRUE, echo  = TRUE}

data(oneiric)
map_params
```

This function will attempt to find parameters for the tinkerbell progression that can be converted into territories. 
Since this is can be a time consuming process, we set a timer for parameter search (in seconds). 
```{r find_tinker, eval = FALSE, echo = TRUE}
map_params <- find_tinkerbell(time = 120,
    n_maps = 96,
    plot = FALSE,
    export = FALSE,
    file_name = "maps")
```

## Preparing output directory

```{r out_dir, eval = TRUE, echo = TRUE}

output <- "Oneiric/simulations/"
```


## Territory Types
We have four territory types you can create:

* Circle: randomly size circles placed in a background territory.
* Rod: randomly size rod-like structures placed in a background territory.
* Chaos: tinkerbell progression chaos map placed in a background.
* Random: will randomly sample one of the above (differing probabilities)



```{r spatial_sim, eval = TRUE, echo = TRUE}
circle <- simulate_spatial(n_cells = 5000,
    n_territories = 5,
    n_samples = 12,
    pattern = "circle",
    expanse = c(0.1, 0.3))

rod <- simulate_spatial(n_cells = 5000,
    n_territories = 5,
    n_samples = 12,
    pattern = "rod",
    width_range = c(0.0, 0.1),
    length_range = c(0.2, 0.5))

chaos_map <- simulate_spatial(n_cells = 5000,
    n_samples = 12,
    pattern = "chaos",
    expanse = 0.02) 

random_map <- simulate_spatial(n_cells = 5000,
    n_territories = 5,
    n_samples = 12,
    pattern = "random",
    width_range = c(0.0, 0.1),
    length_range = c(0.2, 0.5))

```



```{r view_territory_types, eval = TRUE, echo = TRUE}
circles <- do.call("rbind", circle)
circles$Territory <- as.factor(circles$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$Territory)))
g <- ggplot(circles, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


rods <- do.call("rbind", rod)
rods$Territory <- as.factor(rods$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(rods$Territory)))
g1 <- ggplot(rods, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


tinker <- do.call("rbind", chaos_map)
tinker$Territory <- as.factor(tinker$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(tinker$Territory)))
g2 <- ggplot(tinker, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


random <- do.call("rbind", random_map)
random$Territory <- as.factor(random$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(random$Territory)))
g3 <- ggplot(random, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.5) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))

```

```{r g_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Circle Territories"}
print(g)
```

```{r g1_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Rod Territories"}
print(g1)
```

```{r g2_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Chaos Map Territories"}
print(g2)
```

```{r g3_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Random Territories"}
print(g3)
```

# Adding Cell Counts

Once you have created the territories you like, you can add counts to these territories

## Simple Counts

```{r circle_solo, eval = TRUE, echo = TRUE}

circular <- simulate_spatial(n_cells = 5000, 
    n_territories = 5, 
    n_samples = 12, 
    pattern = "circle", 
    expanse = c(0.1, 0.25), 
    force_cells = 10)

circular_counts <- simulate_cells(circular,
    cell_composition = 2,
    no_label = TRUE)

circles <- do.call("rbind", circular_counts$spatial)
circles$Territory <- as.factor(circles$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$Territory)))
g <- ggplot(circles, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.75) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))



circles$cell_labels <- as.factor(circles$cell_labels)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$cell_labels)))
g1 <- ggplot(circles, aes(x = x, y = y, col = cell_labels)) +
    geom_point(size = 0.75) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))

```

```{r no_lab_g_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="No label territories assumes that no cell type information is provided. A single label is given to all cells in the Territory column"}
print(g)
```

```{r no_lab_g1_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="No label territories assumes that no cell type information is provided. Ground truth cell type labels are still provided as the cell_lables column"}
print(g1)
```

In case you would like to export:
```{r export_circle, eval = FALSE, echo =TRUE}
export_simulation(spatial = circular_counts$spatial,
    cells = circular_counts$counts,
    out_dir = output,
    file_tag = "circle_spatial_territories")

```


## Simuating Layer regime

```{r circle_layer_nl, eval = TRUE, echo = TRUE}
circular_layer_nl <- simulate_spatial(n_cells = 5000,
    n_territories = 1,
    n_samples = 12,
    pattern = "circle",
    expanse = c(0.4, 0.5),
    layer = 5,
    force_cells = 12)

circular_layer_counts_nl <- simulate_cells(circular_layer_nl,
    de_prob = 0.5,
    de_layer = 0.05,
    cell_composition = 2,
    no_label = TRUE)

# to mimic no cell type labels available

circles <- do.call("rbind", circular_layer_counts_nl$spatial)
circles$Territory <- as.factor(circles$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$Territory)))
g <- ggplot(circles, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.75) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


circles$cell_labels <- as.factor(circles$cell_labels)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$cell_labels)))
g1 <- ggplot(circles, aes(x = x, y = y, col = cell_labels)) +
    geom_point(size = 0.75) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))




```
```{r no_lab_lay_g_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="No label territories assumes that no cell type information is provided. A single label is given to all cells in the Territory column"}
print(g)
```

```{r no_lab_lay_g1_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="No label territories assumes that no cell type information is provided. Ground truth cell type labels are still provided as the cell_lables column"}
print(g1)
```

In case you would like to export:
```{r export_layer, eval = FALSE, echo = TRUE}
export_simulation(spatial = circular_layer_counts_nl$spatial,
    cells = circular_layer_counts_nl$counts,
    out_dir = output,
    file_tag = "layered_spatial_territories")
```


## Simuating Layer regime with dropped cells

```{r circle_layer_rnd, eval = TRUE, echo = TRUE}
circular_layer_rnd <- simulate_spatial(n_cells = 5000,
    n_territories = 2,
    n_samples = 12,
    pattern = "circle",
    expanse = c(0.15, 0.3),
    layer = 4,
    force_cells = 5)

circular_layer_counts_rnd <- simulate_cells(circular_layer_rnd,
    de_prob = 0.5,
    de_layer = 0.05,
    cell_composition = 2,
    no_label = TRUE,
    randomize_cells = TRUE)

# to mimic no cell type labels available

circles <- do.call("rbind", circular_layer_counts_rnd$spatial)
circles$Territory <- as.factor(circles$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$Territory)))
g <- ggplot(circles, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.75) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


circles$cell_labels <- as.factor(circles$cell_labels)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$cell_labels)))
g1 <- ggplot(circles, aes(x = x, y = y, col = cell_labels)) +
    geom_point(size = 0.75) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))




```
```{r rnd_lay_g_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Dropped cells assumes that not all cells are present between datasets."}
print(g)
```

```{r rnd_lay_g1_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Dropped cells assumes that not all cells are present between datasets."}
print(g1)
```

In case you would like to export:
```{r export_rnd, eval = FALSE, echo = TRUE}
export_simulation(spatial = circular_layer_counts_rnd$spatial,
    cells = circular_layer_counts_rnd$counts,
    out_dir = output,
    file_tag = "dropped_spatial_territories")
```

## Simulating full random data sets 

```{r random_layer_rnd, eval = TRUE, echo = TRUE}
circular_layer_rnd <- simulate_spatial(n_cells = 5000,
    n_territories = 3,
    n_samples = 12,
    pattern = "random",
    expanse = c(0.15, 0.3),
    width_range = c(0, 0.3),
    length_range = c(0.15,0.3),
    layer = seq(1,4),
    force_cells = 5)

circular_layer_counts_rnd <- simulate_cells(circular_layer_rnd,
    de_prob = 0.5,
    de_layer = 0.05,
    cell_composition = 2,
    no_label = TRUE,
    randomize_cells = TRUE)

# to mimic no cell type labels available

circles <- do.call("rbind", circular_layer_counts_rnd$spatial)
circles$Territory <- as.factor(circles$Territory)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$Territory)))
g <- ggplot(circles, aes(x = x, y = y, col = Territory)) +
    geom_point(size = 0.75) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))


circles$cell_labels <- as.factor(circles$cell_labels)
cols <- colorRampPalette(RColorBrewer::brewer.pal(11, "Spectral"))
cols <- cols(length(levels(circles$cell_labels)))
g1 <- ggplot(circles, aes(x = x, y = y, col = cell_labels)) +
    geom_point(size = 0.75) +
    theme_bw() +
    scale_color_manual(values = cols) +
    facet_wrap(~sample) +
    guides(colour = guide_legend(
        override.aes = list(size =  5)))




```
```{r random_lay_g_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Dropped cells assumes that not all cells are present between datasets."}
print(g)
```

```{r random_lay_g1_plot , eval = TRUE, echo = TRUE, fig.width=12,fig.height=11, fig.cap="Dropped cells assumes that not all cells are present between datasets."}
print(g1)
```

